Appendix B

### 2021-05-05 manuscript/bef_stability/appendixB.Rmd
Update appendix following Elisa's review

---
title: Appendix B of Species richness and food-web structure jointly drive community biomass and its temporal stability in fish assemblages 
author: Alain Danet, Maud Mouchet, Willem Bonnaffé, Elisa Thébault and Colin Fontaine
date: \today 
output:
  bookdown::pdf_document2:
    fig_caption: true 
    keep_tex: true
    toc: true
fontsize: 12pt
header-includes:
   - \usepackage{natbib}
   - \newcommand{\beginsupplement}{\setcounter{table}{2}  \renewcommand{\thetable}{S\arabic{table}} \setcounter{figure}{0} \renewcommand{\thefigure}{S\arabic{figure}}} 
   - \usepackage{float}
   - \usepackage{booktabs}
   - \usepackage{makecell}
   - \usepackage{setspace}
   - \doublespacing
   - \usepackage{lineno}
   - \linenumbers
geometry: margin=2cm
bibliography: references.bib
number_sections: true
nocite: |
    @post_ecosystem_2000| @doi_resource_2009| @mchugh_dual_2010|
    @ogorman_unexpected_2017| @gauzens_biodiversity_2020| @dunne_network_2006|
    @mchugh_dual_2010| @doi_resource_2009| @allan_stream_2009|
    @petchey_predicting_2010| @winemiller_must_1989
csl: ecology-letters.csl
---

@@

The monitoring of the fish communities occurs in a variety of stream size, shape
and depth. The stream characteristics constrain the type of protocol that can be
applied. Small streams (depth < 0.7m across the entire width) are sampled with
the complete protocol. The protocol consists of placing anodes across the entire
width of the stream. The fishermen made from one to three rows of electric
fishing. We selected only the data from the first row to homogenize the sampling
effort. In deeper and wider streams (depth >= 0.7m), the sampling was partial
over the bank and the channel and performed by boat and by foot (along the
bank). We selected only the sampling done over the bank because the channel was
not systematically sampled. Two types of partial protocols takes place: partial
over bank and partial by points, which are respectively continuous and discrete.
Furthermore, in the "partial by point" protocol, the channel was sometimes
sampled as well. We selected only the "partial by point" sampling in which all
the sampling points were done over the bank to homogenize the sampling. Then, we
merged partial over bank and partial by point protocols. Some sampled stations
were located in stream with intermediate characteristics where both protocols
were alternatively used depending on the state of the stream at the time of
sampling. For these sites, we only kept the samples that had been carried out
with the most used protocol.

Heterogeneity in the sampling date may introduce variability that is not related
to ecological processes such as the presence of juveniles. In order to reduce
heterogeneity in the sampling date, only one sampling per year was retained if
there was more in a given year, the one with that had the most species and then
the most individuals. Finally, most of the fish sampling occurred in September
(Median: mid-August) and the median standard deviation of the sampling month was
inferior to one (0.7) within stations, i.e. 3 weeks. To further reduce sampling
heterogeneity, we only kept the sampling in which the length of stream sampled
along the stream was inside the median length sampled more or less 30%.

@@

The instructions for body size inference were guided by the recommendations of
the OFB agency, which has defined the sampling protocols, supervised the data
collection and performed analysis for multiple purpose including stream
management.

@@

We inferred the body size of all the unmeasured individuals using the batch
records. There was no need for inference for the N and I batches since the body
size of all individuals was recorded (Fig. \@ref(fig:p-body-size)). The body
size ($l$) of the fishes in the S/L and G was inferred under the assumption that
body size distribution follows a normal law ($l ~ N(\mu, \sigma)$).  For an S/L
batch, the mean $\mu$ and the standard deviation $\sigma$ were simply the
average and the standard deviation of the body size of the subsample recorded.
For a G batch, the average was approximated by $\hat{\mu} = (l_{min} + l_{max})
2^{-1}$ and the standard deviation by $\hat{\sigma} = (l_{min} + l_{max})
4^{-1}$. With the estimations of the mean and the standard deviation, the number
of individuals in a batch and the assumption of normal distribution, we
simulated the body size distribution in S/L and G batches. To avoid simulation
of unrealistic body size, we used a truncated normal distribution bounded at
[0.05; .95] for the density of probability of the S/L batches, and bounded at
[$l_{min}$; $l_{max}$] for the G batches. Body size distribution were generated
using the R function `truncdist::rtrunc`.

Batches that did not meet certain conditions were excluded. First, we also
excluded batches for which the total number of individuals was not recorded. We
also excluded G batches for which $l_{min} \geq l_{max}$ or if
at least one of the measurement was missing. Additionally, a given G batch was
excluded when the number of individuals fell below five, as G batches normally
contained more individuals. We finally excluded S/L batches for which the
number of measured individuals was inferior to 10. Finally, in the dataset used
in the analysis, `r na_length` individuals had missing length over 
`r nrow_fish_length` individuals.

@@

Some species were identified based on their morphotype, but they
belong to the same species based on fecundation criteria. We merged those
morphotype species into one species. There was
`r length(unique(names(sp_to_replace())))` species to merge which belonged to 
`r length(unique(sp_to_replace()))` species.

@@

Diet information were obtained from fishbase.org and literature (Appendix A, Table S2). The
diet information contained from two to three life stages by species, the life
stages being defined by the body size of the fishes. Over the 57 fish species,
only seven had a piscivorous life stage and only three had a strict piscivorous
life stage. Fish species were then largely omnivorous.

@@

#### Define trophic interactions

@@

trophic_building_caption <- paste0( "Building of the metaweb. (1) Each
  species are divided in nine body size classes. We gathered information about
  the food diet of each fish species and how food diet varies according to
  ontogeny. (2) The life stage of a trophic species was determined with the
  median of the body size class. A trophic link was set between the trophic
  species and the resources associated to this life stage. (3) For a piscivorous
  trophic species, the trophic links between fishes were defined according to
  the predators/prey body size ratio. The predation window of the trophic
  species ranged from 3% to 45% of the median of the body size range. A trophic
  link between the predator trophic species and every trophic species whose
  midpoint of bodysize range was included in the predation window.")

@@

```{r, cache = FALSE}
metaweb_caption <- paste0(
  "The metaweb describes the trophic interactions between all the 412 trophic species, each is
  represented by a node and each color represents a species (S = 45). The seven
  resource nodes (7) are drawn in grey. The species names corresponding to the
  three digit codes are found in Table S1."
)
```

@@

The life stage to which a trophic species belongs was determined by the median of
the body size range of the trophic species, each life stage being delimited by a body
size range. In turn, this trophic species has trophic interactions with the
resource nodes of its life stage. The trophic interactions between a
predator and a prey were determined by predator-prey body size ratio
[@brose_predator_2019; @poisot_structure_2016; @schneider_body_2012]. For a
given trophic species, we first determined if its body size range overlap
with a piscivorous life stage (Fig. \@ref(fig:p-net-inf)). We were then more
liberal for piscivory determination than for fish-resource interactions, to
avoid missing potential interactions which were the main interactions in our
community. If the trophic species was piscivorous, we determined its prey size
range from 3% to 45% of the median of its size range. A trophic interaction was
set for each trophic species whose median of its size class fell into the prey
range of the predator trophic species. To sum up, the metaweb contains feeding
interactions (1) between trophic species, (2) between trophic species and
resource nodes and (3) between resource nodes.

#### Local trophic network

The local trophic networks (i.e. one network by
sampling) were obtained by determining the trophic species of each fish
individual in the community. According to the trophic species present in a given
community, we set the trophic interactions between them according to the metaweb. 
All the resource nodes were assumed to be present in each local network.
Finally, the structure of the local networks varied according the presence of
the trophic species.

@@

The procedure to obtain altitude, slope, width and depth of the streams are fully
explained in main text. Because those measurements were collected during the
fishing protocol, they were easier to manage in the analysis. Then, we detailed
in the following sections how water temperature, water flow and Biological
Oxygen Demand (BOD) variables were obtained. The goal was to obtain an annual
mean for each variable, to match the yearly resolution of the fish monitoring.
Furthermore, those variables are measured at different locations than the fish
monitoring.

@@

Water temperature, water flow and BOD variables were obtained from the [naiades
database](http://www.naiades.eaufrance.fr/) and were respectively collected at
an
hourly, daily and monthly frequency. Water temperature and flow are recorded by
automated sensors while BOD is collected manually because its measurement
requires laboratory analysis. BOD data collection is coordinated by the Agence
de l'eau which realized the analysis itself or order it to a company.

```{r, cache = FALSE}
pressure_cap <- paste0(
  "Map of the environmental variables obtained for the 99 sites used in the
  analysis. Slope is expressed in degree, altitude in meter, distance from
  source in kilometer, Strahler order has no units, average stream width and
  depth are expressed in meter, Biological Oxygen Demand (BOD) is expressed in
  mg of dissolved oxygen, water flow is expressed in cube meter per second and
  water temperature is expressed in Celsius degree.")
```

@@

To get rid of seasonal effects of temperature, we applied a moving window
averaging to each time series. The moving window was applied on daily average
temperature. Then we took the annual mean of each year and each variable
(i.e. one value by year and by station).

Several characteristics of the streams make the interpolation more complicated
than along terrestrial surface. Interpolations in terrestrial areas are roughly
based on euclidean distance while euclidean distance is complicated to consider
in streams because it had to follow the path of the streams. Furthermore, two
streams can be close by but not connected and thus the value of their variables might 
not be related or only a few. For those reason, we used stream spatial network
methods to interpolate water temperature, water flow and BOD.

We reconstructed the stream network and the connection between stations with the
`openSTARS` R package [@kattwinkel_preparing_2020]. The reconstruction was done with a raster of digital
elevation model at a resolution of 250 meters [(IGN)](https://wxs.ign.fr/jvam1hsjm11u8voorw81v2xb/telechargement/prepackage/BDALTI-75M_PACK_FXX_2018-01-24%24BDALTIV2_2-0_75M_ASC_LAMB93-IGN69_FRANCE_2018-01-15/file/BDALTIV2_2-0_75M_ASC_LAMB93-IGN69_FRANCE_2018-01-15.7z)
and a theoretical hydrologic network [@pella_reseau_2012].

The interpolations were carried out by hydrographic basin due to convergence
issues when the interpolation was carried out at the scale of the whole
France. We had to find a trade-off between the number of data (to keep enough
high to perform the interpolation) and the level of dataset splitting. We then
merged some hydrographic basin to end up with four regions of interpolation:
West, East, South and North.

Then we modelled each variable as dependant only on the covariation between
sites and taking in account stream network. To model the autocovariance between
sites, we used the linear-sill and mariah functions for taildown and tailup
autocovariance, according to their predictive performance. The interpolations
were done with the `SSN` R package [@isaak_applications_2014; @hoef_ssn:_2014].

@@

cv_dbo_temp_flow %>%
  mutate_if(is.double, ~signif(., 2)) %>%
  kable(
    align = "c",
    format = "latex",
    booktabs = T,
    caption = paste0("Results of the cross-validation for the interpolation of
      BOD, temperature and flow. Average over 4 interpolations are shown. The
      cross-validation were performed with the SSN R package. bias: average
      predicted error, std.biais: bias standardized, RMSPE: Root Mean-Squared
      Prediction Error, std.MSPE: standardized Mean-Squared Prediction Error.") ) %>%
    collapse_rows(columns = c(1, 2), valign = "top") %>%
    kable_styling(latex_options =c("striped", "scale_down"))
```

@@

When predicted values were below the minimum value or above the maximum value
recorded in the dataset for flow, temperature and BOD, we replaced the
predicted value by respectively the minimum and the maximum value found in the
dataset. We ran the interpolation four times, and we kept the predicted value
which had the least standard deviation.

@@

The CV of biomass over time can cover multiple temporal dynamics, including
increases, stable, decreases and a mix of them. Theoretically, however, the CV
computed to quantify stability is often represented with the assumption that the
overall dynamic of the biomass varies around an equilibrium
[@kefi_advancing_2019]. In order to approach the condition of equilibrium, we kept only the
stations that have met this criteria. To characterise the biomass temporal
dynamics, we used the symbolic aggregated approximation method (SAX), which is
commonly used in time series analysis and implemented in the `jmotif` R package
[@senin_grammarviz_2018].  This method classifies time series rather than
quantifies trends and so is generic enough to encompass a diversity of temporal
trajectories. The classification necessitates three steps. First, the time
series are normalized (i.e. substituting the mean and divide by the standard
deviation). The second step is to reduce the dimensionality of the time series
by a Piecewise Aggregation Approximation (PAA) method. Each time series is
divided in $N$ frames of equal size that allows to reduce the dimensionality
of the time series. For each frame, we simply kept the mean value of the time
series. The last step is to qualify each frame into string. Because values of
normalized time series typically follows a normal distribution ($N(0,1)$), a
breakpoint can be computed by dividing the area under the normal distribution
in equal parts. The algorithm attributes an alphabetic letter to each part from
the lowest to the highest ($a$, $b$, $c$, ...). We set three breakpoints, so
$a$, $b$ and $c$.  Finally, the trajectory of each time series is classified as
a combination of $a$, $b$, $c$ letters. $abc$ and $cba$ time series represent
respectively monotonic increasing and decreasing trajectories. For data
analysis, we kept only the $bbb$ which represent time series that showed no
temporal tendencies, with respect to our methodology.

@@

```{r eig-plot}
rotated_eigen_values <- 
  tibble(
    name = paste0("RC", seq(1,14)),
    type = c(rep("Kept", 2), rep("No kept", 14 - 2)),
    value = pca_rotated$rotated$values
  )
eigen_rotated_plot <-
  ggplot(rotated_eigen_values,
    aes(x = reorder(name, -value) , y = value)) +
  geom_col(aes(fill = type)) +
  scale_fill_discrete(name = "Status") +
  labs(x = "Rotated component", y = "Eigen values") +
  theme(legend.position = "none",
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank())
```

```{r contrib-pca, cache = FALSE}
#https://stats.stackexchange.com/a/143949
caption_contrib_pca <- paste0(
  "Loadings of variable to rotated components of the PCA.
  Loadings are the coefficients which determine the components of the PCA as
  linear combinations of the variables.
  "
)

@@

```{r, cache = FALSE}
sem_variable_table <- tibble(
  variable = c(paste0("PCA axis", seq(1, 2)), "Species richness", "Connectance", "Average trophic level", "Synchrony", "CVsp", "Total biomass"),
  unit = c(rep("unitless", 2), "# by square meter", rep("unitless", 4), "gram by square meter"
    ),
  transformation = c("Log base 10", "Log base 10 and inversion", "Log base 10", rep("No", 2), rep("Log base 10", 3)),
  description = c("linked to stream size",
    "linked to average temperature and BOD",
    "Total species richness over the temporal series",
    "Median of annual link density", "Median of annual average trophic level",
    "Temporal synchrony in the dynamics of population biomass", "Average
    variability of population biomass", "Median of annual total biomass")) %>%
kable(
  format = "latex",
  booktabs = T,
  caption = "The description of the variables used in the structural equation models."
  ) %>%
kable_styling(latex_options = c("striped", "scale_down"))
sem_variable_table
```

>>>>

---
title: Appendix B
subtitle: Species richness and food-web structure jointly drive community biomass and its temporal stability in fish assemblages 
author: Alain Danet, Maud Mouchet, Willem Bonnaffé, Elisa Thébault and Colin Fontaine
date: \today 
output:
  bookdown::pdf_document2:
    fig_caption: true 
    keep_tex: true
    toc: true
fontsize: 12pt
header-includes:
   - \usepackage{natbib}
   - \newcommand{\beginsupplement}{\setcounter{table}{2}  \renewcommand{\thetable}{S\arabic{table}} \setcounter{figure}{0} \renewcommand{\thefigure}{S\arabic{figure}}} 
   - \usepackage{float}
   - \usepackage{booktabs}
   - \usepackage{makecell}
   - \usepackage{setspace}
   - \doublespacing
   - \usepackage{lineno}
   - \linenumbers
geometry: margin=2cm
bibliography: references.bib
number_sections: true
nocite: |
    @post_ecosystem_2000| @doi_resource_2009| @mchugh_dual_2010|
    @ogorman_unexpected_2017| @gauzens_biodiversity_2020| @dunne_network_2006|
    @mchugh_dual_2010| @doi_resource_2009| @allan_stream_2009|
    @petchey_predicting_2010| @winemiller_must_1989 | @galiana_spatial_2018 | @ogorman_simple_2019| @oksanen_exploitation_1981| @thebault_trophic_2005| @mccann_more_2009| @jiang_different_2009| @fussmann_ecological_2014| kratina_warming_2012
csl: ecology-letters.csl
---

@@

The monitoring of the fish communities occurs in streams varying in size, shape
and depth. The stream characteristics constrain the type of protocol that can be
applied. Small streams (depth < 0.7m across the entire width) are sampled with
the "complete" protocol. This protocol consists of placing anodes across the
entire width of the stream. As the fishermen made from one to three rows of
electric fishing, we selected only the data from the first row to homogenize the
sampling effort. In deeper and wider streams (depth >= 0.7m), the sampling is
partial over the bank and the channel and it is performed by boat and by foot
(along the bank). We selected only the sampling done over the bank because the
channel was not systematically sampled. Two types of partial protocols takes
place: "partial over bank" and "partial by points", which are respectively
continuous and discrete samplings. We considered both types of partial protocols
in our analysis. In the "partial by point" protocol, the channel was sometimes
sampled as well, so we selected only the "partial by point" sampling in which
all the sampling points were done over the bank to homogenize the sampling. Some
sampled stations were located in streams with intermediate characteristics where
both "complete" and "partial" protocols were alternatively used depending on the
state of the stream at the time of sampling. For these sites, we only kept the
samples that had been carried out with the most used protocol.

Heterogeneity in the sampling date may introduce variability in the fish
communities such as the presence of juveniles. In order to reduce heterogeneity
in the sampling date, only one sampling per year was retained if there were more
in a given year, the one that had the most species and then the most
individuals. Finally, most of the fish sampling occurred in September (Median:
mid-August) and the median standard deviation of the sampling month was inferior
to one (0.7) within stations, i.e. 3 weeks. To further reduce sampling
heterogeneity, we only kept the sampling in which the length of stream sampled
along the stream was inside the median length sampled more or less 30%.

@@

Body size inference was guided by the recommendations of
the OFB agency, which has defined the sampling protocols, supervised the data
collection and performed analyses on the dataset for multiple purpose including stream
management.

@@

We inferred the body size of all the unmeasured individuals using the batch
records. There was no need for inference for the N and I batches since the body
size of all individuals was recorded (Fig. \@ref(fig:p-body-size)). The body
size ($l$) of the fishes in the S/L and G was inferred under the assumption that
body size distribution follows a normal law ($l \sim N(\mu, \sigma)$).  For an S/L
batch, the mean $\mu$ and the standard deviation $\sigma$ were simply the
average and the standard deviation of the body sizes of the subsample recorded.
For a G batch, the average was approximated by $\hat{\mu} = (l_{min} +
l_{max}) / 2$ and the standard deviation by $\hat{\sigma} = (l_{min} + l_{max})
/ 4$. With the estimations of the mean and the standard deviation, the number
of individuals in a batch and the assumption of normal distribution, we
simulated the body size distribution in S/L and G batches. To avoid simulation
of unrealistic body size, we used a truncated normal distribution bounded at
[0.05; .95] for the density of probability of the S/L batches, and bounded at
[$l_{min}$; $l_{max}$] for the G batches. Body size distributions were generated
using the R function `truncdist::rtrunc`.

Batches that did not meet certain conditions were excluded. First, we also
excluded batches for which the total number of individuals was not recorded. We
also excluded G batches for which $l_{min} \geq l_{max}$ or if at least one of
the measurement was missing. Additionally, a given G batch was excluded when the
number of individuals felt below five, as G batches normally contained more
individuals. We finally excluded S/L batches for which the number of measured
individuals was inferior to 10, because it indicates a potential problem in the
batch. Finally, in the dataset used in the analysis, `r na_length` individuals
had missing length over `r nrow_fish_length` individuals, and body size was
inferred for these individuals.

@@

Some species were identified to subspecies, based on their morphotype, but they
belong to the same species based on fecundation criteria. We merged those
subspecies into one species. There was
`r length(unique(names(sp_to_replace())))` species to merge which belonged to 
`r length(unique(sp_to_replace()))` species.

@@

Diet information was obtained from fishbase.org and literature (Appendix A, Table S2). The
diet information contained from two to three life stages by species, the life
stages being defined by the body size of the fishes. Over the 57 fish species,
only seven had a piscivorous life stage and only three had a strict piscivorous
life stage. Fish species were then largely omnivorous.

@@

#### Definition of trophic interactions

@@

trophic_building_caption <- paste0( "Building of the metaweb. (1) Each
  species is divided into nine body size classes. We gathered information about
  the food diet of each fish species and how food diet varies according to
  ontogeny. (2) The life stage of a trophic species is determined by the
  median of the body size class, hereafter midpoint. A trophic link is set between the trophic
  species and the resources associated to this life stage. (3) For piscivorous
  trophic species, the trophic links among fishes are defined according to
  the predator/prey body size ratio. The predation window of a trophic
  species ranges from 3% to 45% of its midpoint. A trophic
  link is set between the predator trophic species and every trophic species whose
  midpoint of body size class range was included in the predation window.")

@@

```{r, cache = FALSE}
metaweb_caption <- paste0(
  "The metaweb describes the trophic interactions between all the 412 trophic species. Each trophic species is
  represented by a node and each color represents a species (S = 45). The seven
  resource nodes are drawn in grey. The species names corresponding to the
  three digit codes are found in Table S1."
)
```

@@

The life stage to which a trophic species belongs was determined by the median
of the body size range of the trophic species, each life stage being delimited
by a body size range. In turn, this trophic species has trophic interactions
with the resource nodes of its life stage. The trophic interactions between a
predator and a prey were determined by the predator-prey body size ratio
[@brose_predator_2019; @poisot_structure_2016; @schneider_body_2012]. For a
given trophic species, we first determined if its body size range overlapped
with a piscivorous life stage (Fig. \@ref(fig:p-net-inf)). If the trophic
species was piscivorous, we determined its prey size range from 3% to 45% of the
median of its size range. A trophic interaction was set for each trophic species
whose corresponding size class median felt into the prey range of the predator trophic
species (Fig. \@ref(fig:p-net-inf)). To sum up, the metaweb contains feeding interactions (1) between
trophic species, (2) between trophic species and resource nodes and (3) between
resource nodes.

#### Local food-webs

The local food-webs (i.e. one network by
sampling) were obtained by determining the trophic species of each fish
individual in the community. According to the trophic species present in a given
community, we set the trophic interactions between them as defined in the metaweb. 
All the resource nodes were assumed to be present in each local network.
Thus, the structure of the local food-webs varied according to the presence of
the trophic species.

@@

The procedures to obtain altitude, slope, width and depth of the streams are fully
explained in main text. Because those measurements were collected during the
fishing protocol, they were easier to manage in the analysis. Then, we detail
in the following sections how water temperature, water flow and Biological
Oxygen Demand (BOD) variables were obtained. The goal was to obtain an annual
mean for each variable, to match the yearly resolution of the fish monitoring.
Furthermore, the available data for these variables were measured at different
locations than the fish monitoring and then necessitates a step of spatial
interpolation.

@@

Water temperature, water flow and BOD variables were obtained from the [naiades
database](http://www.naiades.eaufrance.fr/) and were respectively collected at
an hourly, daily and monthly frequency. Water temperature and flow are recorded
by automated sensors while BOD is collected manually because its measurement
requires laboratory analysis. BOD data collection is coordinated by the Agence
de l'eau (i.e. French national water agency).

```{r, cache = FALSE}
pressure_cap <- paste0(
  "Map of the environmental variables obtained for the 99 sites used in the
  analysis. The inner borders delimit the hydrographic basins. Slope is expressed in degree, altitude in meter, distance from
  source in kilometer, Strahler order has no units, average stream width and
  depth are expressed in meter, Biological Oxygen Demand (BOD) is expressed in
  mg of dissolved oxygen, water flow is expressed in cubic meter per second and
  water temperature is expressed in Celsius degree.")
```

@@

To smooth out the seasonal patterns of temperature variation across year and
keep the general trends, we applied a moving window averaging to each time
series. The moving window was applied on daily average temperature. Then we took
the annual mean of each year and each variable (i.e. one value by year and by
station).

We used stream spatial network methods to interpolate water temperature, water
flow and BOD. Indeed, stream characteristics make interpolation more
complicated than in terrestrial areas. Interpolations in terrestrial areas are
often based on euclidean distance, which is not possible for streams as two
streams can be close but disconnected. For those reason, we used stream spatial
network methods to interpolate water temperature, water flow and BOD.

We reconstructed the stream network and the connection between stations with the
`openSTARS` R package [@kattwinkel_preparing_2020]. The reconstruction was done with a raster of digital
elevation model at a resolution of 250 meters [(link to the raster)](https://wxs.ign.fr/jvam1hsjm11u8voorw81v2xb/telechargement/prepackage/BDALTI-75M_PACK_FXX_2018-01-24%24BDALTIV2_2-0_75M_ASC_LAMB93-IGN69_FRANCE_2018-01-15/file/BDALTIV2_2-0_75M_ASC_LAMB93-IGN69_FRANCE_2018-01-15.7z)
and a theoretical hydrologic network, which is more convenient to use than the observed one [@pella_reseau_2012].

The interpolations were carried out by hydrographic basin due to convergence
issues when the interpolation was carried out at the scale of the whole
France. We had to find a trade-off between the number of data (to keep 
high enough to perform the interpolation) and the level of dataset splitting. We then
merged some hydrographic basins to end up with four regions of interpolation:
West, East, South and North.

We modelled each variable, i.e. water temperature, water flow and BOD, as dependant only on the covariation between
sites taking in account the stream network. To model the autocovariance between
sites, we used the linear-sill and mariah functions for taildown and tailup
autocovariance, according to their predictive performance. The interpolations
were done with the `SSN` R package [@isaak_applications_2014; @hoef_ssn:_2014].

@@

cv_dbo_temp_flow %>%
  mutate_if(is.double, ~signif(., 2)) %>%
  kable(
    align = "c",
    format = "latex",
    booktabs = T,
    caption = paste0("Results of the cross-validation for the interpolation of
      BOD, temperature and flow. The averages over 4 interpolations are shown. The
      cross-validation were performed with the SSN R package. bias: average
      predicted error, std.biais: bias standardized, RMSPE: Root Mean-Squared
      Prediction Error, std.MSPE: standardized Mean-Squared Prediction Error.") ) %>%
    collapse_rows(columns = c(1, 2), valign = "top") %>%
    kable_styling(latex_options =c("striped", "scale_down"))
```

@@

When predicted values were below the minimum value or above the maximum value
recorded in the dataset for flow, temperature and BOD, we replaced the
predicted value by respectively the minimum and the maximum value found in the
dataset. We ran the interpolation four times, and we kept the predicted value
which had the smaller standard deviation.

@@

The CV of biomass over time can cover multiple temporal dynamics, including
increases, stable, decreases and a mix of them. Theoretically, however, the CV
computed to quantify stability is often represented with the assumption that the
overall dynamic of the biomass varies around an equilibrium
[@kefi_advancing_2019]. In order to approach the condition of equilibrium, we
kept only the stations that show no clear temporal trends. To characterise the
biomass temporal dynamics according to their trajectories, we used the symbolic
aggregated approximation method (SAX), which is commonly used in time series
analysis and implemented in the `jmotif` R package [@senin_grammarviz_2018].
This method classifies time series rather than quantifies trends and so is
generic enough to encompass a diversity of temporal trajectories. The
classification necessitates three steps. The time series are first normalized. Then, we reduced the dimensionality of the time series by a Piecewise
Aggregation Approximation (PAA) method. Each time series is divided into $N$
frames of equal size that allows to reduce the dimensionality of the time
series. For each frame, we simply kept the mean value of the time series. The
last step is to qualify each frame into string. Because values of normalized
time series typically follows a normal distribution ($N(0,1)$), a breakpoint can
be computed by dividing the area under the normal distribution in equal parts.
The algorithm attributes an alphabetic letter to each part from the lowest to
the highest ($a$, $b$, $c$, ...). We set three breakpoints, so $a$, $b$ and $c$.
Finally, the trajectory of each time series is classified as a combination of
$a$, $b$, $c$ letters. $abc$ and $cba$ time series represent respectively
monotonic increasing and decreasing trajectories. For data analysis, we kept
only the $bbb$ which represent time series that showed no temporal tendencies,
with respect to our methodology.

@@

```{r eig-plot}
rotated_eigen_values <- 
  tibble(
    name = paste0("RC", seq(1,14)),
    type = c(rep("Kept", 2), rep("No kept", 14 - 2)),
    value = pca_rotated$rotated$values
  )
eigen_rotated_plot <-
  ggplot(rotated_eigen_values,
    aes(x = reorder(name, -value) , y = value)) +
  geom_col(aes(fill = type)) +
  scale_fill_discrete(name = "Status") +
  labs(x = "Rotated component", y = "Eigenvalues") +
  theme(legend.position = "none",
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank())
```

```{r contrib-pca, cache = FALSE}
#https://stats.stackexchange.com/a/143949
caption_contrib_pca <- paste0(
  "Loadings of variable to rotated components of the PCA.
  Loadings are the coordinates of the variables in the space of the PCA
  components.
  "
)

@@

```{r, cache = FALSE, fig.pos = "!H", out.extra = ""}
sem_variable_table <- tibble(
  variable = c(paste0("PCA axis", seq(1, 2)), "Species richness", "Connectance", "Average trophic level", "Synchrony", "CVsp", "Total biomass"),
  unit = c(rep("unitless", 2), "# by square meter", rep("unitless", 4), "gram by square meter"
    ),
  transformation = c("Log base 10", "Log base 10 and inversion", "Log base 10", rep("No", 2), rep("Log base 10", 3)),
  description = c("linked to stream size",
    "linked to average temperature and BOD",
    "Total species richness over the temporal series",
    "Median of annual link density", "Median of annual average trophic level",
    "Temporal synchrony in the dynamics of population biomass", "Average
    variability of population biomass", "Median of annual total biomass")) %>%
kable(
  format = "latex",
  booktabs = T,
  caption = "The description of the variables used in the structural equation models."
  ) %>%
kable_styling(latex_options = c("striped", "scale_down"))
sem_variable_table
```

\begin{table}
\caption{\label{tab:hyp-sem} Presentation of the links 
of the Structural Equation Models associated with the references supporting the links.}
\begin{center}
\begin{tabular}{ m{7cm}| m{1cm}| m{6cm}}
    \hline
    Link & sign & reference\\
    \hline
    PCA1 Avg stream size $\to$ Species richness & (+) & Galiana et al. (2018) \\
    \hline
    PCA2 Avg temperature $\to$ Species richness & (+/-) & Woods et al. (2020); Gauzens et al. (2020) \\
    \hline
    PCA1 Avg stream size $\to$ Connectance & (-) & Galiana et al. (2018) \\
    \hline
    PCA2 Avg temperature $\to$ Connectance & (+/-) & O’Gorman et al. (2019);
    Oksanen et al. (1981)\\
    \hline
    Species richness $\to$ Connectance & (-) & Winemiller (1989); Dunne (2006) \\
    \hline
    PCA1 Avg stream size $\to$ Avg trophic level & (+) & Post et al. (2000); Doi et al. (2009); McHugh et al. (2010) \\
    \hline
    PCA2 Avg temperature $\to$ Avg trophic level & (+/-) & O’Gorman et al. (2017)\\
    \hline
    Species richness $\to$ Avg trophic level & (+) & Schneider et al. (2016)\\
    \hline
    Species richness $\to$ Synchrony & (-) & Thébault and Loreau (2005); Loreau and de Mazancourt (2008); Thibaut and Connolly (2013) \\
    \hline
    Connectance $\to$ Synchrony & (+) & Thébault and Loreau (2005)\\
    \hline
    Avg trophic level $\to$ Synchrony & (+) & Raimondo et al. (2004); McCann and Rooney (2009) \\
    \hline
    PCA1 Avg stream size $\to$ Synchrony & (-) & Wang et al. (2019)\\
    \hline
    PCA2 Avg temperature $\to$ Synchrony & (?) & \\
    \hline
    Species richness $\to$ CVsp & (+) & Jiang et al. (2009)\\
    \hline
    Connectance $\to$ CVsp & (+) & Thébault and Loreau (2005)\\
    \hline
    Avg trophic level $\to$ CVsp & (-) & McCann and Rooney (2009); Shanafelt and Loreau (2018)\\
    \hline
    PCA1 Avg stream size $\to$ CVsp & (?) & \\
    \hline
    PCA2 Avg temperature $\to$ CVsp & (+) & Kratina et al. (2012); Fussmann et al. (2014)\\
    \hline
    CVsp $\to$ Biomass stability & (-) & Thibaut and Connolly (2013)\\
    \hline
    Synchrony $\to$ Biomass stability & (-) & Thibaut and Connolly (2013)\\
    \hline
    Species richness $\to$ Total biomass & (+) & Schneider et al. (2016)\\
    \hline
    Connectance $\to$ Total biomass & (+/-) & Thébault & Loreau (2003); Poisot
    et al. (2013)\\
    \hline
    Avg trophic level $\to$ Total biomass & (+) & Wang and Brose (2018)\\
    \hline
    PCA1 Avg stream size $\to$ Total biomass & (+) & Post et al. (2000); Allan and Ibanez Castillo (2009); Doi et al. (2009)\\
    \hline
    PCA2 Avg temperature $\to$ Total biomass & (+/-) & Kratina et al. (2012);
    Tabi et al. (2019)\\
    \hline
    \end{tabular}
\end{center}
\end{table}
